---
title: MTH6139 Time Series
subtitle: Coursework 1 
author: Andrew Samet
date: 19/3/26 
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab 
    highlight: tango
---
 
```{r, echo=FALSE}
# This code will display the QMUL logo at the top right of the page
# Do not change this code
htmltools::img(src = knitr::image_uri("images/QMlogo.png"),
               alt = 'logo',
               style = 'position:absolute; top:0; right:0; padding:10px; width:20%;')
```

# Section 1: My Interests
What interest me is the trend in the results of football matches for my favourite team. 

## Section 1.1: Why This Interest Me?
This interest me because I am big football fan, more specfically a big Man United fan.
Despite the fact that it is very difficult to model football results well as they are so volatile,
it is still a really interesting exercise to create time series and see how the model predicts 
Man United results for the future.

# Section 2: My Time Series
I have a large dataset with football results for European's top five leagues between 2020 and 2021. In my time series the y values are 1/(the best betting odds for each home game), this gives us a good indication for the probability of a home win. There are 19 data points per season, since each team plays 19 home matches a season.

## Section 2.1: Time Series
This is a time series for Man United

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(prophet)
library(arrow)
dataset <- read_parquet("football.parquet")
dataset_filtered <- dataset[dataset$HomeTeam=="Man United",]
df <- data.frame('ds'=dataset_filtered$Date, 'y'=1/dataset_filtered$BbMxH)
m <- prophet(df)
future <- make_future_dataframe(m,100)
forecast = predict(m,future)
dyplot.prophet(m,forecast)
prophet_plot_components(m, forecast)
```

## Section 2.2: Observations
It is very interesting to see in the "weekly" plot that around Thursday the probability drop to -0.05 week on week. This is reflected in the fact that not much football is played on Thursdays. Therefore, it makes a lot of sense to see this decrease.

# Section 3: Another Time Series
This is another time series for Man City. I thought it would be interesting to compare the two, firstly, because Man City are big rivals of my team, Man United. Secondly, because Man City's success has taken a very different path to that of Man United over the past 20 years.

## Section 3.1: Time Series
```{r echo=FALSE, message=FALSE, warning=FALSE}
dataset_filtered <-dataset[dataset$HomeTeam=="Man City",]
df <- data.frame('ds'=dataset_filtered$Date, 'y'=1/dataset_filtered$BbMxH)
m <- prophet(df)
future <- make_future_dataframe(m,100)
forecast = predict(m,future)
dyplot.prophet(m,forecast)
prophet_plot_components(m, forecast)
```

## Section 3.3: Observations
Again we see this dip on the weekly trend on Thursdays.

# Section 4: Summary
Perhaps the most striking thing to note is the gradual rise in success for Man City vs the gradual decline in success for Man United. You can see in the 'trend' plot, that Man City's success has increased healthily over time, whereas Man United's has decreased. Furthermore, one can observe from the interactive plots that the model predicts continued success for Man City over the coming 100 games, whereas, for Man United the future doesn't look quite as exciting. 